Low-Thrust Trajectory Design Near Asteroids Using Reachability Sets

### 2016-08-11 manuscript.tex
Add some corrections from Christine

 \author{
  Shankar Kulumani\thanksibid{1}%
    \thanks{Doctoral Student, \href{mailto:skulumani@gwu.edu}{skulumani@gwu.edu}. Student AIAA Member.}
  \ and Taeyoung Lee\thanksibid{2}\thanks{Associate Professor, \href{mailto:tylee@gwu.edu}{tylee@gwu.edu}. AIAA Member}\\
  {\normalsize\itshape
   George Washington University, Washington, District of Columbia }\\
   }

@@

% Motivation for missions/studying asteroids
Small solar system bodies, such as asteroids and comets, are of significant interest to the scientific community.
These small bodies offer great insight into the early formation of the solar system.
This insight offers additional detail into the formation of the Earth and also the probable formation of other planetary bodies.
Of particular interest are those near-Earth asteroids (NEA) which inhabit heliocentric orbits in vicinity of the Earth.
These easily accessible bodies provide attractive targets to support space industrialization, mining operations, and scientific missions..
NEAs potentially contain many materials useful for propulsion, construction, semiconductors, precious and strategic metals~\cite{ross2001}.
In addition, these NEAs are also of concern for their potential to impact the Earth~\cite{wie2008}.
Asteroids and comets are the greatest threat to future civilizations and as a result there is a focused effort to mitigate these risks.
In spite of the great interest in asteroids, the operation of spacecraft in their vicinity is a challenging problem.

% Difficulty in system model
While there has been significant study of interplanetary transfer trajectories, relatively less analysis has been conducted on operations in the vicinity of asteroids.
The dynamic environment around asteroids is strongly perturbed and challenging for analysis and mission operation.
Due to their small size, resulting in a low gravitational attraction, asteroids have irregular shapes or potentially chaotic spin states.
Furthermore, due to their low mass other non-gravitational effects, such as solar radiation pressure or third-body effects, become much more significant.
As a result, the orbital environment is in general quite complex and it is difficult to generate analytical insights.

@@

% use of low thrust propulsion
The application of optimal control methods for orbital trajectory design is nontrivial.
In order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the optimal solution.
However, the asteroid system dynamics are nonlinear and exhibit chaotic behaviors.
This makes solving the optimization problem highly dependent on the initial condition.
Similar to the three-body problem, there is an insufficient number of analytical constants to derive an analytical solution in general.
As a result, accurate numerical methods are required to determine optimal solutions.
These methods are critically dependent on accurate initial guess in order to allow for convergence to an optimal solution.

% similarity to the three body problem
The irregular shape of asteroids leads to a richer dynamics as compared to the familiar Kepler's problem.
The motion of particles around asteroids is modeled using the restricted full two-body problem.
This model has many similarities with the restricted three body problem, and much of the theory developed in the three-body problem is also applicable.
The similarity of the three-body problem to motion around asteroids allows for much of the same theory to be applied to the asteroid problem~\cite{mondelo2010,herrera2014}.
In this paper, we extend the design method developed in the three-body problem to motion about asteroids~\cite{kulumani2015}.
Our approach avoids the difficulties in selecting an appropriate initial guess for optimization.
Instead, we utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft.
We calculate the reachable set on a lower dimensional \Poincare section which allows a 
In addition, the generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer.
This simple methodology allows for extended transfer trajectories which iteratively approach a desired target orbit.

@@

In this analysis we consider transfers about the asteroid 4769 Castalia.
Castalia has an accurate shape model and is also considered a potentially hazardous asteroid with a possibility of Earth impact~\cite{hudson1994}.
We model the gravitational potential field of Castalia using a polyhedron gravity model instead of using a spherical harmonic expansion.
The spherical harmonic expansion is a popular method of representing the gravity field~\cite{scheeres1996}.
Approximations are possible by truncating the infinite order series to fixed set of coefficients, with the most important terms corresponding to the second order and degree~\cite{scheeres1994}.
The approximation error is generally small given a sufficient amount of terms. 
However, when evaluated close to the body the series expansion will diverge and is no longer accurate.

@@

% benefits of the polyhedron model
A polyhedral model of the surface of an asteroid can be determined from remote optical or radar sensors.
The faces of the polyhedron can be large or small and allows for fine detail such as depression, craters, ridges, or interior voids.
In addition, there is no requirement for the body to be modeled at a uniformly high resolution so small details can be incorporated with minimal cost.
From the shape model, an analytical, closed form expression for the gravitational potential can be derived.
The polyhedral approach provides an accurate gravitational model consistent with the resolution of the shape and the chosen discretization.
Furthermore, the polyhedron model is an exact solution up to the surface of the body.
Therefore, this model is ideal for missions traversing large regions both close and far from the asteroid.

@@

We represent the gravitational potential of the asteroid using a polyhedron gravitation model.
This model is composed of a polyhedron, a three-dimensional solid body, which is defined by a series of vectors in the body-fixed frame.
The vectors define vertices in the body-fixed frame as well as planar faces which compose the surface of the asteroid.
We assume that each face is a triangle composed of three vertices and three edges.
As a result, only two faces meet at each edge while three faces meet at each vertex.
Only the body-fixed vectors, and their associated topology, is required to define the exterior gravitational model.
References~\citenum{werner1994} and~\citenum{werner1997} give a detailed derivation of the polyhedron model.
Here, we summarize the key developments and equations required for implementation.

@@

One interesting thing to note is that both~\cref{eq:face_dyad,eq:edge_dyad} can be precomputed without knowledge of the position of the satellite.
They are both solely functions of the vertices and edges of the polyhedral shape model and are computed once and stored.
Once a position vector \( \vecbf{r} \) is defined then the scalars given in~\cref{eq:edge_factor,eq:face_factor} can be computed for each face and edge.
Finally,~\cref{eq:potential} is used to compute the gravitational potential on the spacecraft.
The Laplacian, defined in~\cref{eq:laplacian}, gives a simple method to determine if the spacecraft has collided with the body~\cite{werner1996}

In this work, we consider trajectories about asteroid 4769 Castalia.
Doppler radar images, obtained at the Arecibo Observatory in 1989, are used to determine a shape model of Castalia~\cite{hudson1994,neese2004}.
We use the estimated rotation period of \SI{4.07}{\hour} with a nominal density of \SI{2.1}{\gram\per\centi\meter\cubed}~\cite{scheeres1996}.
The shape model is composed of \num{4092} triangular faces and a rendering of the asteroid is provided in~\cref{fig:castalia_3d}.
In addition, we also show a contour plot of the radius of Castlia in~\cref{fig:radius_contour}.
\begin{figure}
    \centering
    \begin{subfigure}[htbp]{0.45\textwidth}
        \includegraphics[width=\textwidth]{castalia}
        \caption{3D Shape Model of Castalia} \label{fig:castalia_3d}
    \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[htbp]{0.45\textwidth}
        \includegraphics[width=\textwidth]{radius_contour}
        \caption{Radius contours of Castalia} \label{fig:radius_contour}
    \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
    \caption{Polyhedron Shape Model of 4769 Castalia}
    \label{fig:castalia}
\end{figure}

@@

Typical optimal control methods, including both indirect and direct based methods, are highly dependent on an accurate initial guess.
For indirect optimization, which is based on the calculus of variations, this results in the well-known two-point boundary value problem.
Insight into the problem or insight by the designer is usually required to determine appropriate initial costates that will converge to the optimal solution and satisfy the desired constraints.
To avoid this issue, we utilize the concept of the reachability set on a lower dimensional \Poincare section.
By repeatedly constructing the reachability set we can achieve general transfers by determining set intersections on the \Poincare section.
This alleviates the need to determine an accurate initial guess while offering some insight into the dynamics of neighboring trajectories

@@

We seek to approximate the reachability set on a \Poincare section by solving an optimal control problem.
The \Poincare section is chosen in a manner similar to the previous work in both the three-body problem as well analysis performed around asteroids to determine periodic orbits.
Analysis in the three-body problem relies heavily on symmetries in the force fields.
However, in our system model the gravitational potential, given by~\cref{eq:potential}, has no symmetries.
In spite of this, it is still possible to determine periodic solutions through the application of a \Poincare map with surface of section chosen normal to a surface in the phase space~\cite{scheeres2000}.
For a periodic orbit, the trajectories will intersect the \Poincare section at two distinct points every half orbit.
With the addition of a low thrust control input, we are able to expand the reachable set from a distinct point to a larger area on the \Poincare section.
\Cref{fig:reachability_set} illustrates this methodology.
Without any control input, the trajectories will intersect the \Poincare section at \( \vecbf{x}_n\).
The addition of a control input allows the spacecraft to depart from the natural dynamics and intersect the section at another location.
We use the cost function \( J \) to define a distance metric on the \Poincare section.
Maximization of \( J \), or the minimization of \( -J \), along various directions on the \Poincare section allows us to generate the largest reachability set under the bounded control input.

@@

An optimal control problem is defined by the cost function
\begin{align}\label{eq:cost}
    J = -\frac{1}{2} \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right)^T 
    Q
    \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right) ,
\end{align}
where \( \vecbf{x}_n(t_f) \) is the final state of a control-free trajectory, while the term \( \vecbf{x}(t_f)\) is the final state of a trajectory under the influence of the control input.
We use the matrix \( Q = \text{diag} \begin{pmatrix} 1 & 0 & 1 & 1 & 0 &1 \end{pmatrix} \in \R^{6 \times 6}\) to represent the mapping onto \( \Sigma \).
Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eq:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
We ensure that the trajectories intersect the \Poincare section through the use of terminal constraints.
In addition, we use the terminal constraints to define a specific direction along which we seek to minimize the cost~\cref{eq:cost}.
Since the \Poincare section is four-dimensiona, we parameterize a direction in \( \R^4 \)  using three angles \( \phi_1, \phi_2 , \phi_3 \).
The terminal constraints are given as
\begin{align}\label{eq:terminal_constraints}
    \begin{split}
        m_1 &= y = 0 , \\
        m_2 &= \parenth{\sin \phi_{1_{d}}} \parenth{ x_1^2 + x_2^2 + x_3^2 + x_4^2} - x_1^2 = 0, \\
        m_3 &= \parenth{\sin \phi_{2_{d}}} \parenth{ x_2^2 + x_3^2 + x_4^2} - x_2^2 = 0, \\
        m_4 &= \parenth{\sin \phi_{3_{d}}} \parenth{ 2 x_3^2 + 2 x_3 \sqrt{x_4^2 + 2 x_4^2}} - x_3 - \sqrt{x_4^2 + x_3^2} = 0 ,
    \end{split}
\end{align}
where we make use of the difference states \( \parenth{x_1, x_2 ,x_3, x_4 }\) defined as
\begin{align}\label{eq:diff_states}
    \begin{split}
        x_1 &= x(t_f) - x_n(t_f) , \\
        x_2 &= z(t_f) - z_n(t_f) , \\
        x_3 &= \dot{x}(t_f) - \dot{x}_n(t_f) , \\
        x_4 &= \dot{z}(t_f) - \dot{z}_n(t_f) . \\
    \end{split}
\end{align}
The constraint \( m_1 = 0 \) ensures that the terminal states lies on the \Poincare section.
The constraints \( m_2, m_3, m_4 \) are used to define a direction on the \Poincare section.
\Cref{eq:diff_states} represents the difference between our controlled and uncontrolled trajectory on the \Poincare section.
We approximate the entire reachable set by discretization  over the space of angles \(\phi_1, \phi_2, \phi_3 \).
By convention we assume that the angles lie in the following range
\begin{align*}
    \phi_1, \phi_2 &\in [ 0, \pi ) ,\\
    \phi_3 &\in [ 0 , 2 \pi ) ,
\end{align*}
such that we parameterize all directions on the three sphere, \(\S^3\).
Finally, we also incorporate a control acceleration magnitude constraint given by
\begin{align}\label{eq:control_constraint}
    c(\vecbf{u}) = \vecbf{u}^T \vecbf{u} - u_m^2 \leq 0 ,
\end{align}
where \( u_m \) is the maximum acceleration possible by the propulsion system.
This constraint assumes that the control acceleration may be orientated in any direction yet the acceleration magnitude is variable but bounded.
The goal is to determine the control history \( \vecbf{u}(t) \) such that the cost function~\cref{eq:cost} is minimized while subject to the equations of motion~\cref{eq:body_eoms} and the constraints~\cref{eq:control_constraint,eq:terminal_constraints}.

% How do you solve the optimal control problem
We apply a standard calculus of variations approach to solve our optimal control problem~\cite{bryson2002}.
Using the Euler-Lagrange equations we arrive at the necessary conditions for optimality
\begin{align}\label{eq:necc_conditions}
    \begin{split}
        \dot{\vecbf{x}} ^T &= \deriv{H}{\vecbf{\lambda}} ,\\
        \dot{\vecbf{\lambda}}^T &= \deriv{H}{\vecbf{x}} , \\
        0 &= \deriv{\phi}{x}^T + \deriv{\vecbf{m}}{x}^T \vecbf{\beta} - \vecbf{\lambda}^T(t_f) , \\
        0 &= \deriv{H}{\vecbf{u}} + \mu^T \deriv{c}{\vecbf{u}} ,
    \end{split}
\end{align}
where the Hamiltonian, \( H\), is defined as
\begin{align}\label{eq:hamiltonian}
    H = \vecbf{\lambda}_r^T \vecbf{v} + \vecbf{\lambda}_v^T \parenth{\vecbf{g}(\vecbf{r}) + \vecbf{h}(\vecbf{v}) + \vecbf{u}}.
\end{align}
The costate is given by \( \vecbf{\lambda} = \begin{bmatrix} \vecbf{\lambda}_r & \vecbf{\lambda}_v \end{bmatrix}^T \in \R^{6 \times 1}\), \( \vecbf{\beta} \in \R^{4 \times 1} \) are the additional Lagrange multiplers associated with the terminal constraints in~\cref{eq:terminal_constraints}, and \( \mu \) is a Lagrange multipler associated with the control constraint in~\cref{eq:control_constraint}.

@@

With the reachability set encompassing the target orbit, we can generate a final transfer to the target.
We use the minimum state calculated from the final stage to serve as the initial condition of the transfer.
A final optimal transfer is computed to satisfy the fixed terminal state \( \vecbf{x}(t_f) = \vecbf{x}_t \) and the bounded control magnitude constraint.
\Cref{fig:trajectory} shows the entire transfer trajectory, from the three reachable set trajectories as well as the final transfer to the target.
\begin{figure}[htbp] 
    \centering 
    \begin{subfigure}[htbp]{0.45\textwidth} 
        \includegraphics[width=\textwidth]{figures/trajectory.pdf} 
        \caption{Equatorial view of transfer} \label{fig:trajectory_up} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.45\textwidth} 
        \includegraphics[width=\textwidth]{figures/trajectory_3d.pdf} 
        \caption{Out of plane view} \label{fig:trajectory_3d} 
    \end{subfigure}~ 
    \caption{Complete transfer trajectory}
    \label{fig:trajectory} 
\end{figure}
It is interesting to note the while both the initial and target periodic orbit lie in the equatorial plane, the reachability trajectories show a relatively large out of plane component during the transfers.
In spite of this out of plane movement, the reachability set approaches and meets the target orbit. 
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/control.pdf}
    \caption{Control history \label{fig:control}}
\end{figure}
\Cref{fig:control} shows the control input required during the maneuver.
The transfers show a large out of plane movement for a planar transfer
We can see that the control constraint in~\cref{eq:control_constraint} is satisfied over the entire trajectory.

>>>>

 \author{
  Shankar Kulumani\thanksibid{1}%
    \thanks{Doctoral Student, \href{mailto:skulumani@gwu.edu}{skulumani@gwu.edu}. Student AIAA Member.}
  \ and Taeyoung Lee\thanksibid{2}\thanks{Associate Professor, \href{mailto:tylee@gwu.edu}{tylee@gwu.edu}. AIAA Member.}\\
  {\normalsize\itshape
   George Washington University, Washington, District of Columbia }\\
   }

@@

% Motivation for missions/studying asteroids
Small solar system bodies, such as asteroids and comets, are of significant interest to the scientific community; these small bodies offer great insight into the early formation of the solar system.
This insight offers additional detail into the formation of the Earth and also the probable formation of other planetary bodies.
Of particular interest are those near-Earth asteroids (NEA) which inhabit heliocentric orbits in vicinity of the Earth.
These easily accessible bodies provide attractive targets to support space industrialization, mining operations, and scientific missions.
NEAs potentially contain many materials such as those useful for propulsion, construction, or for use in semiconductors.
Also, many bodies contain highly profitable materials, such as precious or strategic metals~\cite{ross2001}.
In addition, these NEAs are also of concern for their potential to impact the Earth.
Asteroids and comets are the greatest threat to future civilizations and as a result there is a focused effort to mitigate these risks~\cite{wie2008}.
In spite of the great interest in asteroids, the operation of spacecraft in their vicinity is a challenging problem.

% Difficulty in system model
While there has been significant study of interplanetary transfer trajectories, relatively less analysis has been conducted on operations in the vicinity of asteroids.
The dynamic environment around asteroids is strongly perturbed and challenging for analysis and mission operation.
Due to their low mass, which results in a low gravitational attraction, asteroids may have irregular shapes and potentially chaotic spin states.
Furthermore, since the magnitude of the gravitational attraction is relatively small non-gravitational effects, such as solar radiation pressure or third-body effects, become much more significant.
As a result, the orbital environment is generally quite complex and it is difficult to generate analytical insights.

@@

% use of low thrust propulsion
The application of optimal control methods for orbital trajectory design is nontrivial.
Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the optimal solution.
However, the asteroid system dynamics are nonlinear and exhibit chaotic behaviors.
This makes solving the optimization problem highly dependent on the initial condition.
Similar to the three-body problem, there is an insufficient number of analytical constants to derive an analytical solution in general.
As a result, accurate numerical methods are required to determine optimal solutions.
These methods are critically dependent on accurate initial guess in order to allow for convergence to an optimal solution.

% similarity to the three body problem
The irregular shape of asteroids leads to a richer dynamics as compared to the familiar Kepler's problem.
The motion of particles around asteroids is modeled using the restricted full two-body problem.
This model has many similarities with the restricted three body problem, and much of the theory developed for the three-body problem is also applicable.
The similarity of the three-body problem to motion around asteroids allows for much of the same theory to be applied to the asteroid problem~\cite{mondelo2010,herrera2014}.
In this paper, we extend the design method developed in the three-body problem to motion about asteroids~\cite{kulumani2015}.
Our approach avoids the difficulties in selecting an appropriate initial guess for optimization.
We instead utilize the concept of the reachability set to enable a simple methodology of selecting initial conditions to achieve general orbital transfers.
Given an initial condition and fixed time horizon, the reachable set is the set of states attainable, subject to the operational constraints of the spacecraft.
We calculate the reachable set on a lower dimensional \Poincare section. 
The generation of the reachable set allows for a more systematic method of determining initial conditions and eases the burden on the designer.
This simple methodology allows for extended transfer trajectories which iteratively approach a desired target orbit.

@@

In this analysis we consider transfers about the asteroid 4769 Castalia.
Castalia has an accurate shape model and is also considered a potentially hazardous asteroid with a possibility of Earth impact~\cite{hudson1994}.
We model the gravitational potential field of Castalia using a polyhedron gravity model instead of using a spherical harmonic expansion.
The spherical harmonic expansion is a popular method of representing the gravity field~\cite{scheeres1996}.
Approximations are possible by truncating the infinite order series to fixed set of coefficients, with the most important terms corresponding to the second order and degree~\cite{scheeres1994}. 
However, when evaluated close to the body the series expansion will diverge and is no longer accurate. 
Therefore, the spherical harmonic representation is not ideal for landing trajectories or those passing close to the surface.

@@

% benefits of the polyhedron model
A polyhedral model of the surface of an asteroid can be determined from remote optical or radar sensors.
The faces of the polyhedron can be large or small and allow for fine detail such as depression, craters, ridges, or interior voids.
In addition, there is no requirement for the body to be modeled at a uniformly high resolution so small details can be incorporated with minimal cost.
From the shape model, an analytical, closed form expression for the gravitational potential can be derived.
The polyhedral approach provides an accurate gravitational model consistent with the resolution of the shape and the chosen discretization.
Furthermore, the polyhedron model is an exact solution up to the surface of the body.
Therefore, this model is ideal for missions traversing large regions both close and far from the asteroid.

@@

We represent the gravitational potential of the asteroid using a polyhedron gravitation model.
This model is composed of a polyhedron, a three-dimensional solid body, that is defined by a series of vectors in the body-fixed frame.
The vectors define vertices in the body-fixed frame as well as planar faces which compose the surface of the asteroid.
We assume that each face is a triangle composed of three vertices and three edges.
As a result, only two faces meet at each edge while three faces meet at each vertex.
Only the body-fixed vectors, and their associated topology, is required to define the exterior gravitational model.
References~\citenum{werner1994} and~\citenum{werner1996} give a detailed derivation of the polyhedron model.
Here, we summarize the key developments and equations required for implementation.

@@

One interesting thing to note is that both~\cref{eq:face_dyad,eq:edge_dyad} can be precomputed without knowledge of the position of the satellite.
They are both solely functions of the vertices and edges of the polyhedral shape model and are computed once and stored.
Once a position vector \( \vecbf{r} \) is defined, the scalars given in~\cref{eq:edge_factor,eq:face_factor} can be computed for each face and edge.
Finally,~\cref{eq:potential} is used to compute the gravitational potential on the spacecraft.
The Laplacian, defined in~\cref{eq:laplacian}, gives a simple method to determine if the spacecraft has collided with the body~\cite{werner1996}

In this work, we consider trajectories about asteroid 4769 Castalia.
Doppler radar images, obtained at the Arecibo Observatory in 1989, are used to determine a shape model of Castalia~\cite{hudson1994,neese2004}.
We use the estimated rotation period of \SI{4.07}{\hour} with a nominal density of \SI{2.1}{\gram\per\centi\meter\cubed}~\cite{scheeres1996}.
The shape model is composed of \num{4092} triangular faces and a rendering of the asteroid is provided in~\cref{fig:castalia_3d}.
In addition, we show a contour plot of the radius of Castlia in~\cref{fig:radius_contour}.
\begin{figure}
    \centering
    \begin{subfigure}[htbp]{0.45\textwidth}
        \includegraphics[width=\textwidth]{castalia}
        \caption{3D Shape Model of Castalia} \label{fig:castalia_3d}
    \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[htbp]{0.45\textwidth}
        \includegraphics[width=\textwidth]{radius_contour}
        \caption{Radius contours of Castalia} \label{fig:radius_contour}
    \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)
    \caption{Polyhedron Shape Model of 4769 Castalia}
    \label{fig:castalia}
\end{figure}

@@

Typical optimal control methods, including both indirect and direct based methods, are highly dependent on an accurate initial guess.
For indirect optimization, which is based on the calculus of variations, this results in the well-known two-point boundary value problem.
Insight into the problem or insight by the designer is usually required to determine appropriate initial costates that will converge to the optimal solution and satisfy the desired constraints.
To avoid this issue, we utilize the concept of the reachability set on a lower dimensional \Poincare section.
By repeatedly constructing the reachability set, we can achieve general transfers by determining set intersections on the \Poincare section.
This alleviates the need to determine an accurate initial guess while offering some insight into the dynamics of neighboring trajectories

@@

We seek to approximate the reachability set on a \Poincare section by solving an optimal control problem.
The \Poincare section is chosen in a manner similar to the previous work in both the three-body problem as well analysis performed around asteroids to determine periodic orbits.
Typically, analysis for the three-body problem relies heavily on symmetries in the force fields.
However, in our system model, the gravitational potential, given by~\cref{eq:potential}, has no symmetries.
In spite of this, it is still possible to determine periodic solutions through the application of a \Poincare map with the surface of section chosen normal to a surface in the phase space~\cite{scheeres2000}.
For a periodic orbit, the trajectories will intersect the \Poincare section at two distinct points every half orbit.
With the addition of a low thrust control input, we are able to expand the reachable set from a distinct point to a larger area on the \Poincare section.
\Cref{fig:reachability_set} illustrates this methodology.
Without any control input, the trajectories will intersect the \Poincare section at \( \vecbf{x}_n\).
The addition of a control input allows the spacecraft to depart from the natural dynamics and intersect the section at another location.
We use the cost function \( J \) to define a distance metric on the \Poincare section.
Maximization of \( J \), or the minimization of \( -J \), along various directions on the \Poincare section allows us to generate the largest reachability set under the bounded control input.

@@

An optimal control problem is defined by the cost function
\begin{align}\label{eq:cost}
    J = -\frac{1}{2} \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right)^T 
    Q
    \left( \vecbf{x}(t_f) - \vecbf{x}_{n}(t_f)\right) ,
\end{align}
where \( \vecbf{x}_n(t_f) \) is the final state of a control-free trajectory, while the term \( \vecbf{x}(t_f)\) is the final state of a trajectory under the influence of the control input.
We use the matrix \( Q = \text{diag} \begin{pmatrix} 1 & 0 & 1 & 1 & 0 &1 \end{pmatrix} \in \R^{6 \times 6}\) to represent the mapping onto \( \Sigma \).
Maximization of the distance between \( \vecbf{x}_n \) and \(\vecbf{x} \), on the \Poincare section defined in~\cref{eq:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
We ensure that the trajectories intersect the \Poincare section through the use of terminal constraints.
In addition, we use the terminal constraints to define a specific direction along which we seek to minimize the cost~\cref{eq:cost}.
Since the \Poincare section is four-dimensiona, we parameterize a direction in \( \R^4 \)  using three angles \( \phi_1, \phi_2 , \phi_3 \).
The terminal constraints are given as
\begin{align}\label{eq:terminal_constraints}
    \begin{split}
        m_1 &= y = 0 , \\
        m_2 &= \parenth{\sin \phi_{1_{d}}} \parenth{ x_1^2 + x_2^2 + x_3^2 + x_4^2} - x_1^2 = 0, \\
        m_3 &= \parenth{\sin \phi_{2_{d}}} \parenth{ x_2^2 + x_3^2 + x_4^2} - x_2^2 = 0, \\
        m_4 &= \parenth{\sin \phi_{3_{d}}} \parenth{ 2 x_3^2 + 2 x_3 \sqrt{x_4^2 + 2 x_4^2}} - x_3 - \sqrt{x_4^2 + x_3^2} = 0 ,
    \end{split}
\end{align}
where we make use of the difference states \( \parenth{x_1, x_2 ,x_3, x_4 }\) defined as
\begin{align}\label{eq:diff_states}
    \begin{split}
        x_1 &= x(t_f) - x_n(t_f) , \\
        x_2 &= z(t_f) - z_n(t_f) , \\
        x_3 &= \dot{x}(t_f) - \dot{x}_n(t_f) , \\
        x_4 &= \dot{z}(t_f) - \dot{z}_n(t_f) . \\
    \end{split}
\end{align}
The constraint \( m_1 = 0 \) ensures that the terminal state lies on the \Poincare section.
The constraints \( m_2, m_3, m_4 \) are used to define a direction on the \Poincare section.
\Cref{eq:diff_states} represents the difference between our controlled and uncontrolled trajectory on the \Poincare section.
We approximate the entire reachable set by discretization  over the space of angles \(\phi_1, \phi_2, \phi_3 \).
By convention we assume that the angles lie in the following range
\begin{align*}
    \phi_1, \phi_2 &\in [ 0, \pi ) ,\\
    \phi_3 &\in [ 0 , 2 \pi ) ,
\end{align*}
such that we parameterize all directions on the three sphere, \(\S^3\).
Finally, we also incorporate a control acceleration magnitude constraint given by
\begin{align}\label{eq:control_constraint}
    c(\vecbf{u}) = \vecbf{u}^T \vecbf{u} - u_m^2 \leq 0 ,
\end{align}
where \( u_m \) is the maximum acceleration possible by the propulsion system.
This constraint assumes that the control acceleration may be orientated in any direction yet the acceleration magnitude is variable but bounded.
The goal is to determine the control history \( \vecbf{u}(t) \) such that the cost function~\cref{eq:cost} is minimized while subject to the equations of motion~\cref{eq:body_eoms} and the constraints~\cref{eq:control_constraint,eq:terminal_constraints}.

% How do you solve the optimal control problem
We apply a standard calculus of variations approach to solve our optimal control problem~\cite{bryson1975}.
Using the Euler-Lagrange equations we arrive at the necessary conditions for optimality
\begin{align}\label{eq:necc_conditions}
    \begin{split}
        \dot{\vecbf{x}} ^T &= \deriv{H}{\vecbf{\lambda}} ,\\
        \dot{\vecbf{\lambda}}^T &= \deriv{H}{\vecbf{x}} , \\
        0 &= \deriv{\phi}{x}^T + \deriv{\vecbf{m}}{x}^T \vecbf{\beta} - \vecbf{\lambda}^T(t_f) , \\
        0 &= \deriv{H}{\vecbf{u}} + \mu^T \deriv{c}{\vecbf{u}} ,
    \end{split}
\end{align}
where the Hamiltonian, \( H\), is defined as
\begin{align}\label{eq:hamiltonian}
    H = \vecbf{\lambda}_r^T \vecbf{v} + \vecbf{\lambda}_v^T \parenth{\vecbf{g}(\vecbf{r}) + \vecbf{h}(\vecbf{v}) + \vecbf{u}}.
\end{align}
The costate is given by \( \vecbf{\lambda} = \begin{bmatrix} \vecbf{\lambda}_r & \vecbf{\lambda}_v \end{bmatrix}^T \in \R^{6 \times 1}\), \( \vecbf{\beta} \in \R^{4 \times 1} \) are the additional Lagrange multiplers associated with the terminal constraints in~\cref{eq:terminal_constraints}, and \( \mu \) is a Lagrange multipler associated with the control constraint in~\cref{eq:control_constraint}.

@@

With the reachability set encompassing the target orbit, we can generate a final transfer to the target.
We use the minimum state calculated from the final stage to serve as the initial condition of the transfer.
A final optimal transfer is computed to satisfy the fixed terminal state \( \vecbf{x}(t_f) = \vecbf{x}_t \) and the bounded control magnitude constraint.
\Cref{fig:trajectory} shows the entire transfer trajectory, from the three reachable set trajectories as well as the final transfer to the target.
\begin{figure}[htbp] 
    \centering 
    \begin{subfigure}[htbp]{0.45\textwidth} 
        \includegraphics[width=\textwidth]{figures/trajectory.pdf} 
        \caption{Equatorial view of transfer} \label{fig:trajectory_up} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.45\textwidth} 
        \includegraphics[width=\textwidth]{figures/trajectory_3d.pdf} 
        \caption{Out of plane view} \label{fig:trajectory_3d} 
    \end{subfigure}~ 
    \caption{Complete transfer trajectory}
    \label{fig:trajectory} 
\end{figure}
It is interesting to note that while both the initial and target periodic orbit lie in the equatorial plane, the reachability trajectories show a relatively large out of plane component during the transfers.
In spite of this out of plane movement, the reachability set approaches and meets the target orbit. 
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/control.pdf}
    \caption{Control history \label{fig:control}}
\end{figure}
\Cref{fig:control} shows the control input required during the maneuver.
The transfers show a large out of plane movement for a planar transfer
We can see that the control constraint in~\cref{eq:control_constraint} is satisfied over the entire trajectory.
